Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta

### 2024-07-31 tex/main.tex
add consideration of coverage and response time objectives

@@

\textcolor{red}{Note that in Equations~\ref{eqn:survival_A}
and~\ref{eqn:survival_B}, the survival functions used, $s_k(t)$, are the same
whether a primary or a secondary vehicle arrives first. This may be unrealistic,
where primary vehicles may be better equipped than secondary vehicles to deal
with some emergencies. One option to overcome this would be to have distinct
survival functions for each vehicle type, however there may be some interactions
between the survival probability functions that would not be captured, for
example when a primary vehicle arrives after a secondary vehicle, there may
still be a better chance of survival that the secondary vehicle only. This would
be an interesting avenue of further research. For this work, survival functions
will be the same for all vehicles, representing the probability of surviving
given time to be seen by any emergency vehicle.}

@@

We propose finding the true vehicle utilisations by first solving
Equation~\ref{eqn:utilisations_primary} for the $\lambda_a$; determining the
primary vehicle utilisations with Equation~\ref{eqn:utilisation_ratio_primary};
then using these to solve Equation~\ref{eqn:utilisations_secondary} for the
$\tilde{\lambda}_a$; and determining the secondary vehicle utilisations with
Equation~\ref{eqn:utilisation_ratio_secondary}. These can be solved
numerically. In our implementation this is solved using the MINPACK hybrd and
hybrj algorithms by using Scipy's \texttt{fsolve} function
\citep{2020SciPy-NMeth}. \textcolor{red}{These algorithms are efficient methods for finding zeros of systems of non-linear equations. They are based on the Powell hybrid method \cite{powell70}.} It should be noted that in this section we assume that
all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
that will be used throughout the optimisation methodology, however in the
simulation we can account for time-dependent demand.

@@

\subsection{Planning for Future Demand}\label{sec:demand_scenarios}
Similarly to many other cities worldwide, ambulance services in Jakarta
anticipate that demand for services will grow in future years, not least because
there is now a coordinated service accessible via a single common number `119'
to call. This should help raise awareness and visibility of an ambulance
service. In the case of Jakarta, certainly increased use of 119 and therefore
increased demand would be a desired outcome and is pro-actively supported
by the Indonesian Government.

@@

\textcolor{red}{
\subsection{Comparing Expected \& Simulated Survival Measures}
In Table~\ref{tbl:current_optimal_compare} it can be seen that there is a large
discrepancy between the expected survival, both overall and of A1 patients,
computed using the MESLMHPHF described in Section~\ref{sec:meslmhphf} and
calculated using the simulation. This is a consequence of the variability in
response times, which is explicitly modelled by the simulation but not captured
by the MESLMHPHF. As seen in the Table, the standard deviations of the response
times are relatively large in comparison to the mean response times, and so this
effect would be substantial here.}

\textcolor{red}{
The discrepancies are exacerbated by the use of cut-off survival function curves
for most of the patients: consider a survival function with cut-off $L$, with
response time below $L$ the patient's chance of survival is 0, and above $L$ the
patient's chance of survival is guaranteed. The optimisation will find a
solution such that will ensure that the expected response time is above $L$,
possibly as low as $L + \epsilon$, but will not attempt to push the expected
response time above this, as it will have no effect on the expected survival.
However considering variability of mean response times, a considerable
proportion of the response times will lie below the expected value, and so
possibly below $L$, resulting in worse survival probabilities, as observed in
Table~\ref{tbl:current_optimal_compare}. Therefore the distributions chosen to
model travel times can have a large impact on the survival measures, in this
case the naive Triangular distribution.}

\textcolor{red}{
This leads to a number of avenues of future work to improve this and overcome
this effect. Variability could be explicitly modelled within the MESLMHPHF, such
that it is not only maximising \textit{expected} survival, but considers some
percentiles of guaranteed survival. Another avenue would be to use the
simulation explicitly in the optimisation process, incorporating the simulation
into the objective function and using simulation-optimisation techniques to
improve runtime efficiency. However, as discussed in Section~\ref{sec:survival},
maximising expected survival of all patients can lead to a fairer allocation of
resources, with difficult to reach patients being given equal consideration to
closer, easier to reach patients. When considering the variability of travel
times in the optimisation process, difficult to reach patients may be ignored, indistinguishable from patients who happen to sample a response time at the
extreme tail of a given distribution curve.}

\textcolor{red}{
Therefore, due to this interpretation of fairness of allocations, and due to the
naive choice of travel time distributions, including difficulties in being able
to validate such a model and the choices made, we consider only the MESLMHPHF
objective from now on.
}

@@

A key feature of both models, and their novelty, is the consideration of
heterogeneous fleets. In a highly populated area such as Jakarta (around
16,000 population per km$^2$), the inclusion of RRVs such as paramedics on
motorbikes is very important, given RRVs can access areas that cannot easily
or quickly be accessed by ambulances. Results from both the optimisation and
simulation research showed that the RRVs can help increase the overall
survival and reduce the response times, especially crucial when responding to
life threatening events such as cardiac arrests \citep{holmen2020shortening}.
\textcolor{red}{In building the models, some assumptions or relaxations were
made to simplify the model, and future work should investigate the effects of
these, for example the assumption that service times are not location dependent
is unrealistic, but relaxed here. Further developments of the model could
consider explicitly modelling service rates using the travel time matrices.}

>>>>

\textcolor{red}{Finding allocations that maximise survival can result in
allocations with different behaviours to those found by minimising response
times or maximising coverage. Coverage maximisation models are often easier to
solve, and easier to communicate, but may not take into consideration severity
or intensity of demand \citep{Erkut200842}. This results in a minority of
patients, particularly those more isolated, being over-serviced, that is, having
more relative resources allocated to their needs that the majority, at the
expense of the majority. Minimising average response times on the other hand
treats all patients equally, but may have the unintended outcome of disregarding
some more further out patients whose response times would be larger that others
regardless of the allocation. It would still disregard the severity or type of
the patient. Allocations based on maximising survival are similar to those that
minimise response time, overcoming the issue of over-servicing outlying
patients, though it is still possible to disregard them completely. It has an
advantage over minimising response times as it can account for patient severity.
Having survival as functions of response times has the effect of weighting gains
in response times by how valuable those gains are to a patient.}

@@

\textcolor{red}{Note that in Equations~\ref{eqn:survival_A}
and~\ref{eqn:survival_B}, the survival functions used, $s_k(t)$, are the same
whether a primary or a secondary vehicle arrives first. This may be unrealistic,
where primary vehicles may be better equipped than secondary vehicles to deal
with some emergencies. One option to overcome this would be to have distinct
survival functions for each vehicle type, however there may be some interactions
between the survival probability functions that would not be captured, for
example when a primary vehicle arrives after a secondary vehicle, there may
still be a better chance of survival that the secondary vehicle only. This would
be an interesting avenue of further research. For this work, survival functions
will be the same for all vehicles, representing the probability of surviving
given time to be seen by any emergency vehicle.
Note also that the survival functions are functions of $b_{pa}$, the expected
travel time between $p$ and $a$, and do not consider the variability in the
travel times. This will be discussed further alongside results in
Section~\ref{sec:compare_exp_and_sim}.}

@@

We propose finding the true vehicle utilisations by first solving
Equation~\ref{eqn:utilisations_primary} for the $\lambda_a$; determining the
primary vehicle utilisations with Equation~\ref{eqn:utilisation_ratio_primary};
then using these to solve Equation~\ref{eqn:utilisations_secondary} for the
$\tilde{\lambda}_a$; and determining the secondary vehicle utilisations with
Equation~\ref{eqn:utilisation_ratio_secondary}. These can be solved
numerically. In our implementation this is solved using the MINPACK hybrd and
hybrj algorithms by using Scipy's \texttt{fsolve} function
\citep{2020SciPy-NMeth}. \textcolor{red}{These algorithms are efficient methods for finding zeros of systems of non-linear equations. They are based on the Powell hybrid method \citep{powell70}.} It should be noted that in this section we assume that
all $\lambda_{pk}$ are static and not time-dependant, which is an assumption
that will be used throughout the optimisation methodology, however in the
simulation we can account for time-dependent demand.

@@

\subsection{Demand Scenarios}\label{sec:demand_scenarios}
Similarly to many other cities worldwide, ambulance services in Jakarta
anticipate that demand for services will grow in future years, not least because
there is now a coordinated service accessible via a single common number `119'
to call. This should help raise awareness and visibility of an ambulance
service. In the case of Jakarta, certainly increased use of 119 and therefore
increased demand would be a desired outcome and is pro-actively supported
by the Indonesian Government.

@@

\textcolor{red}{
\subsection{Comparing Expected \& Simulated Survival Measures}\label{sec:compare_exp_and_sim}
In Table~\ref{tbl:current_optimal_compare} it can be seen that there is a large
discrepancy between the expected survival, both overall and of A1 patients,
computed using the MESLMHPHF described in Section~\ref{sec:meslmhphf} and
calculated using the simulation. This is a consequence of the variability in
response times, which is explicitly modelled by the simulation but not captured
by the MESLMHPHF. As seen in the table, the standard deviations of the response
times are relatively large in comparison to the mean response times, and so this
effect would be substantial here.}

\textcolor{red}{
The discrepancies are exacerbated by the use of cut-off survival function curves
for most of the patients: consider a survival function with cut-off $L$, with
response time below $L$ the patient's chance of survival is 0, and above $L$ the
patient's chance of survival is guaranteed. The optimisation will find a
solution aim to push the expected response time above $L$, and once this is
reached, possibly with an expected response time as low as $L$ itself, it will
not attempt to push the expected
response time above this, as it will have no effect on the expected survival.
However considering variability of mean response times, a considerable
proportion of the response times will lie below the expected value, and so
possibly below $L$, resulting in worse survival probabilities, as observed in
Table~\ref{tbl:current_optimal_compare}. Therefore the distributions chosen to
model travel times can have a large impact on the survival measures, in this
case the naive Triangular distribution.}

\textcolor{red}{
This leads to a number of avenues of future work to improve this and overcome
this effect. Variability could be explicitly modelled within the MESLMHPHF, such
that it is not only maximising \textit{expected} survival, but considers some
percentiles of guaranteed survival. Another avenue would be to use the
simulation explicitly in the optimisation process, incorporating the simulation
into the objective function and using simulation-optimisation techniques to
improve runtime efficiency.}

\textcolor{red}{
Due to the naive choice of travel time distributions, including difficulties in
being able to validate such a model and the choices made, we consider only the
MESLMHPHF objective from now on.
}

@@

A key feature of both models, and their novelty, is the consideration of
heterogeneous fleets. In a highly populated area such as Jakarta (around
16,000 population per km$^2$), the inclusion of RRVs such as paramedics on
motorbikes is very important, given RRVs can access areas that cannot easily
or quickly be accessed by ambulances. Results from both the optimisation and
simulation research showed that the RRVs can help increase the overall
survival and reduce the response times, especially crucial when responding to
life threatening events such as cardiac arrests \citep{holmen2020shortening}.

\textcolor{red}{In building the models, some assumptions or relaxations were
made to simplify the model, and future work should investigate the effects of
these, for example the assumption that service times are not location dependent
is unrealistic, but relaxed here. Further developments of the model could
consider explicitly modelling service rates using the travel time matrices.
Another limitation identified was the lack of consideration of variability in
travel times in the optimisation model, which cased some discrepancies between
the simulated and expected survivals. One method to overcome this could be to
incorporate variability directly into the MESLMHPHF objective, for example using
percentiles instead of averages, or to optimise using the simulation as an
objective. However in both approaches, a good estimate of travel time
distributions would be needed, with better validation.
}
